My title

### 2024-11-04 paper/paper.qmd
MODEL VISUALZATION

# Turn end_date into quantifiable column, to use for model
harrisdata <- harrisdata %>%
  mutate(
    end_date_num = as.numeric(end_date - min(end_date)))

harrisdata <- harrisdata |>
  mutate(state = factor(state))

@@

bayesian_model <- stan_glmer(
  formula = model,
  data = harrisdata,
  family = binomial(link = "logit"),
  prior = priors,
  prior_intercept = priors,
  weights = harrisdata$numeric_grade,
  seed = 123,
  cores = 4,
  adapt_delta = 0.95)
```

@@

This is Model 2

spline_model <- stan_glm(
  pct ~ ns(end_date_num, df = 5) + pollster + state, 
  data = harrisdata,
  family = gaussian(),
  prior = normal(0, 5),
  prior_intercept = normal(50, 10),
  seed = 1234,
  iter = 2000,
  chains = 4,
  refresh = 0
)

# Summarize the model
summary(spline_model)

# Posterior predictive checks
pp_check(spline_model)

modelsummary::modelsummary(
  list(
    "First model" = spline_model
  ),
  statistic = "mad",
  fmt = 2
)
```

@@

harrisdata$pollster <- factor(harrisdata$pollster)
harrisdata$state <- factor(harrisdata$state)

new_data <- data.frame(
  end_date_num = seq(
    min(harrisdata$end_date_num),
    max(harrisdata$end_date_num),
    length.out = 100
  ),
  pollster = factor("YouGov", levels = levels(harrisdata$pollster)),
  state = factor("Georgia", levels = levels(harrisdata$state)))

posterior_preds <- posterior_predict(spline_model, newdata = new_data)

pred_summary <- new_data |>
  mutate(
    pred_mean = colMeans(posterior_preds),
    pred_lower = apply(posterior_preds, 2, quantile, probs = 0.025),
    pred_upper = apply(posterior_preds, 2, quantile, probs = 0.975)
  )

# Plot data and Bayesian model predictions with confidence ribbon
ggplot(harrisdata, aes(x = end_date_num, y = pct, color = state)) +
  geom_point() +
  geom_line(
    data = pred_summary,
    aes(x = end_date_num, y = pred_mean),
    color = "blue",
    inherit.aes = FALSE
  ) +
  geom_ribbon(
    data = pred_summary,
    aes(x = end_date_num, ymin = pred_lower, ymax = pred_upper),
    alpha = 0.2,
    inherit.aes = FALSE
  ) +
  labs(
    x = "Days since earliest poll",
    y = "Percentage",
    title = "Poll Percentage over Time"
  ) +
  facet_wrap(~ state, scales = "free_y") +
  theme_minimal()
```

# TRUMP #

```{r}
trumpdata <- preddata %>%
  filter(state %in% c("Georgia", "Arizona", "Nevada", "Michigan", "North Carolina", "Pennsylvania", "Wisconsin")) %>%
  filter(candidate_name == "Donald Trump") %>%
  mutate(end_date = mdy(end_date)) %>%
  filter(end_date >= as.Date("2024-07-21")) %>%
  mutate(num_trump = round((pct / 100) * sample_size, 0))

ggplot(trumpdata, aes(x = reorder(state, pct, median), y = pct, fill = state)) +
  geom_boxplot() +
  coord_flip() +  # Flip for better readability
  labs(x = "State", y = "Support Percentage") +
  theme_minimal() +
  theme(legend.position = "none")

trumpdata <- trumpdata %>%
  mutate(
    end_date_num = as.numeric(end_date - min(end_date)))

trumpdata <- trumpdata |>
  mutate(state = factor(state))

# Define Model
model2 <- cbind(num_trump, sample_size - num_trump) ~ (1 | state) + (1 | pollster) + (1 | end_date_num)

bayesian_model2 <- stan_glmer(
  formula = model2,
  data = trumpdata,
  family = binomial(link = "logit"),
  prior = priors,
  prior_intercept = priors,
  weights = trumpdata$numeric_grade,
  seed = 123,
  cores = 4,
  adapt_delta = 0.95)
```

# Plot checks for Model
pp_check(bayesian_model2)
plot(bayesian_model2, pars = "(Intercept)", prob = 0.95)
modelsummary::modelsummary(
  list(
    "First model" = bayesian_model2
  ),
  statistic = "mad",
  fmt = 2
)
```

@@

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

```{r}
harrisdata <- harrisdata %>% filter(!is.na(numeric_grade)) %>% filter(!is.na(num_harris))
trumpdata <- trumpdata %>% filter(!is.na(numeric_grade)) %>% filter(!is.na(num_trump))
trumpdata
```

```{r}
library(dplyr)
library(ggplot2)
library(rstanarm)

harris_data <- harrisdata %>%
  mutate(
    candidate_name = "Kamala Harris",
    support_ratio = num_harris  # Assuming harris_support_ratio is already calculated
  ) %>%
  select(end_date_num, candidate_name, support_ratio)

# Prepare Trump data with a unified structure
trump_data <- trumpdata %>%
  mutate(
    candidate_name = "Donald Trump",
    support_ratio = num_trump  # Assuming trump_support_ratio is already calculated
  ) %>%
  select(end_date_num, candidate_name, support_ratio)

# Combine both into a single data frame
combined_data <- bind_rows(harris_data, trump_data)

library(ggplot2)
library(ggplot2)

# Assuming `combined_data` contains columns `end_date_num`, `support_ratio`, and `candidate_name`

ggplot(combined_data, aes(x = end_date_num, y = support_ratio, color = candidate_name, fill = candidate_name)) +
  # Shaded area around each line for overlap visualization
  geom_ribbon(data = combined_data %>% filter(candidate_name == "Kamala Harris"),
              aes(ymin = support_ratio - 0.05, ymax = support_ratio + 0.05),
              alpha = 0.3, fill = "red") +
  geom_ribbon(data = combined_data %>% filter(candidate_name == "Donald Trump"),
              aes(ymin = support_ratio - 0.05, ymax = support_ratio + 0.05),
              alpha = 0.3, fill = "blue") +

  # Line plot for each candidate
  geom_line(aes(group = candidate_name), size = 1) +

  scale_color_manual(values = c("Kamala Harris" = "red", "Donald Trump" = "blue")) +  # Set line colors
  labs(
    x = "Days Since Earliest Poll", 
    y = "Support Ratio",
    color = "Candidate",
    title = "Support Ratio for Harris and Trump Over Time with Overlap Highlighted"
  ) +
  theme_minimal() +
  theme(
    text = element_text(size = 12),
    plot.title = element_text(hjust = 0.5),
    legend.position = "top",
    legend.title = element_text(face = "bold")
  )

```
# Discussion

>>>>

# Turn end_date into quantifiable column, to use for model
harrisdata <- harrisdata %>%
  mutate(
    end_date_num = as.numeric(end_date - min(end_date)),
        state = factor(state))

@@

bayesian_model <- stan_glmer(
  formula = model,
  data = harrisdata,
  family = binomial(link = "logit"),
  prior = priors,
  prior_intercept = priors,
  seed = 123,
  cores = 4,
  adapt_delta = 0.95)
```

@@

@@

harrisdata <- harrisdata %>% drop_na()
expected_probs <- posterior_epred(bayesian_model, newdata = harrisdata)
median_probs <- apply(expected_probs, 2, median)
harrisdata <- harrisdata %>%
  mutate(median_prob = median_probs)

ggplot(harrisdata, aes(x = end_date_num, y = median_prob, color = state)) +
  geom_line(alpha = 0.5) +  # Plot raw data with some transparency to show noise
  geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs", k = 10), se = FALSE) +  # Smoother with flexible spline
  labs(
    title = "Predicted Probability of Harris Outcome Over Time by State",
    x = "Days Since First Poll (end_date_num)",
    y = "Predicted Probability"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~ state, scales = "free_y")

ggplot(harrisdata, aes(x = end_date_num, y = median_prob, color = state)) +
  geom_line(alpha = 0.5) +  # Plot raw data with some transparency
  geom_smooth(method = "loess", span = 0.3, se = FALSE) +  # Add a less smoothed trend line
  labs(
    title = "Predicted Probability of Harris Outcome Over Time by State",
    x = "Days Since First Poll (end_date_num)",
    y = "Predicted Probability"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~ state, scales = "free_y")

@@

# Discussion
